An investigation of the clinical utility of preoperative cardiopulmonary exercise testing in patients undergoing major pancreatic surgery.

### 2016-06-17 Presentations/viva_presentation_short.tex

\begin{frame}
	\frametitle{Cardiopulmonary exercise testing}
	\framesubtitle{What, how, why?}
	\begin{itemize}

		\item Objective measure of cardiopulmonary response to exercise
		\item Dynamic test measuring multiple aspects of oxygen delivery
		\item Cardiac, respiratory, vascular and mitochondrial function
		\item First described for preoperative assessment by Older and co-workers
		\item Widely used in general, oesophago-gastric, vascular and thoracic surgery
		\item CPET Methodology
		\begin{itemize}
			\item Cycle ergometer with electronic braking
			\item Breath-by-breath gas analysis using a tight-fitting face mask
			\item Continuous monitoring of 12-lead ECG
			\item Initial 3-minute rest period on bike for baseline parameters
			\item Incremental work-load test to volitional fatigue
		\end{itemize}
	\end{itemize}
\end{frame}

@@

\begin{frame}
	\frametitle{Aims of Thesis}
	\begin{enumerate}
		\item To evaluate the clinical utility of preoperative CPET in predicting postoperative adverse events after pancreaticoduodenectomy.
		\item To examine the patient factors that are related to cardiopulmonary exercise physiology with particular attention to the effect of obstructive jaundice and body composition.
		\item To examine the effect of preoperative systemic inflammation and poor aerobic capacity on the magnitude of the post-operative systemic inflammatory response.
		\item To examine the value of serial daily postoperative markers of systemic inflammatory response in the prediction of post-operative complications.
	\end{enumerate}
\end{frame}

%--CHAPTER 2----CHAPTER 2----CHAPTER 2----CHAPTER 2----CHAPTER 2----CHAPTER 2----CHAPTER 2--
\section[Chapter 2]{CPET and Postoperative Outcomes}
\begin{frame}
	\frametitle{CPET and Postoperative Outcomes}
	\framesubtitle{Does CPET have a role in risk stratification of patients?}
	\begin{block}{\textbf{Complications}}
		Low $\dot{V}_{O_2}$AT associated with pancreatic fistula (16\% vs 35\%) \\
		Not associated with cardio-respiratory complications or mortality
	\end{block}

	\begin{block}{\textbf{Length of stay}}
		Low $\dot{V}_{O_2}$AT associated with longer hospital stay (14 vs 20 days)\\
		Not associated with critical care stay or admissions
	\end{block}

	\begin{block}{\textbf{Adjuvant Therapy}}
		Low $\dot{V}_{O_2}$AT associated with non-progression to adjuvant therapy after surgery.
	\end{block}
\end{frame}

@@

		\column{0.6\textwidth}
			{\footnotesize 	Why is $\dot{V}_{O_2}$AT so low in patients undergoing pancreatic surgery?\\
				\medskip
				What factors are associated with a low $\dot{V}_{O_2}$AT?\\
				\medskip
				Why was $\dot{V}_{O_2}$AT not associated with cardio-respiratory complications?\\
				\medskip
				Does $\dot{V}_{O_2}$AT truly reflect cardiopulmonary function in this cohort of patients?\\	
			}

@@

		\begin{figure}
			\includegraphics[width=\textwidth]{jaundice_vs_patient_factors}
		\end{figure}

		\column[c]{0.5\textwidth}		
		{\scriptsize			
			Severity of systemic inflammation proportionate to severity of jaundice. \\
			\medskip
			Obstructive jaundice associated with electrolyte abnormalities.\\
			\medskip
			Jaundiced patients were more likely to be anaemic, with lower haematocrit and lower mean corpuscular volume.\\
			\medskip
			No significant difference in the preoperative renal function, prothrombin time and white cell count between jaundiced and non-jaundiced patients.\\
		}
	\end{columns}
\end{frame}

\begin{frame}
	\frametitle{What preop factors were associated with low $\dot{V}_{O_2}$?}
	\framesubtitle{Multivariate binary logistic regression analysis}
	Factors associated with $\dot{V}_{O_2}$AT $<$ 10 ml/kg/min
	\begin{table}
		\begin{tabular}{l| l l l}
			& HR   & 95\% CI    & p     \\ \hline
			Female sex & 3.75 & 1.57-8.95  & 0.003 \\
			BMI$>$25   & 3.65 & 1.61-8.26  & 0.002 \\
			Cancer     & 4.02 & 1.33-12.16 & 0.014 \\
			CRP$>$10   & 2.98 & 1.29-6.86  & 0.010
		\end{tabular}
	\end{table}
	\hrule
	\medskip
	Factors associated with $\dot{V}_{O_2}$Peak $<$ 16ml/kg/min
	\begin{table}
		\begin{tabular}{l| l l l}
			& HR   & 95\% CI   & p        \\ \hline
			Female sex  & 7.57 & 3.09-18.5 & $<$0.001 \\
			BMI$>$25    & 2.57 & 1.18-5.63 & 0.018    \\
			Haemoglobin & 2.43 & 1.05-5.63 & 0.038
		\end{tabular}
	\end{table}
\end{frame}

@@

	\vfill

	For example, if the absolute $\dot{V}_{O_2}$AT was 0.84 litres/min, the corrected $\dot{V}_{O_2}$AT (ml/kg/min) would vary with patient weight as follows:

@@

@@

	\begin{figure}
		\includegraphics[height=0.4\textheight]{../Figures/bc_scatter_VO2_TAT}
	\end{figure}
	\begin{figure}
		\includegraphics[height=0.45\textheight]{../Figures/bc_scatter_VO2_TAT_elbm}
	\end{figure}

@@

\begin{frame}
	\frametitle{Impact of perioperative systemic inflammation?}
	\framesubtitle{Inflammation before and after surgery affects short and long term outcomes}
	\begin{itemize}
		\item Perioperative systemic inflammation influences short and long-term outcomes
		\item Elevated preoperative systemic inflammation is associated with postoperative complications in colorectal surgery, after oesophagectomy, liver resections and cardiac surgery. 
		\item Early exaggerated postoperative systemic inflammatory response is also associated with increased infective complications
		\item In patients undergoing pancreatic surgery, preoperative inflammatory state is further influenced by:
		\begin{itemize}
			\item Obstructive jaundice
			\item Cholangitis
			\item ERCP and stenting
			\item Post-ERCP / EUS related complications
		\end{itemize}
		\item Obesity is associated with chronic inflammation
		\item Impact of aerobic capacity on postoperative inflammation not studied before
	\end{itemize}
\end{frame}

@@

\section{Case study}
\begin{frame}
	\frametitle{A case study}
	\framesubtitle{How does this work influence decision-making?}
	\begin{itemize}
		\item 63 yr old female with BMI of 35 presents with painless OJ
		\item PMH - HTN, T2DM, ex-smoker
		\item Bloods: Bil 180, CRP 22, Alb 30
		\item Imaging: Resectable tumour of head of pancreas
		\medskip

		\item Should this patient undergo CPET?
		\item Should this patient undergo biliary drainage?
		\item Should we wait until resolution of jaundice before CPET?
		\item How can the CPET results be used to influence her care?
		\item How can the postoperative inflammatory markers be used to manage her care?
	\end{itemize}
\end{frame}

\begin{frame}
	\frametitle{A case study}
	\begin{description}
	{\scriptsize 	
		\item [Ch 2] CPET may allow identifying if she is at increased risk of complications, prolonged hospital stay and non-receipt of adjuvant therapy. 
					\\ CPET should not be used on its own to decline surgery. 
					\hrule
		\item [Ch 3] CPET can be performed in the presence of jaundice. 
					\\ Jaundice does not adversely affect dynamic cardiopulmonary response to and this cannot be a justification for biliary drainage. 
					\\ Taken together with evidence from trials on preperative biliary drainage, she should proceed direct to surgery.
						\hrule
		\item [Ch 4] Interpret low $\dot{V}_{O_2}$ in her case with caution. 
					\\ Her aerobic capacity may be better than that suggested by her test results. 
					\\ Beware of \textit{spurious correlation} with total body weight.
						\hrule
		\item [Ch 5] Jaundice and hypo-albuminemia attenuate post-op inflammatory response.
					\\ However, this does not translate into increased complications. 
					\\ Aerobic capacity does not affect post-op inflammatory response.
						\hrule
		\item [Ch 6] If her drain amylase on D3 is normal and D3 CRP$<$178 or D4 CRP$<$125, she is likely to have an uneventful postoperative course and must continue on 'enhanced-recovery' pathway.
				\\ If she develops a POPF, the early inflammatory response does not predict outcome.
	}
	\end{description}
\end{frame}

>>>>

\begin{frame}
	\frametitle{Cardiopulmonary exercise testing}
	\framesubtitle{What, how, why?}
	{\footnotesize
	\begin{itemize}
		\item Objective measure of cardiopulmonary response to exercise
		\item Dynamic test measuring multiple aspects of oxygen delivery
		\item Cardiac, respiratory, vascular and mitochondrial function
		\item First described for preoperative assessment by Older and co-workers
		\item Widely used in general, oesophago-gastric, vascular and thoracic surgery
	\end{itemize}
	\vfill

	\begin{columns}
		\column[c]{0.6\textwidth}
			\textbf{CPET Methodology}
			\\ - Cycle ergometer with electronic braking
			\\ - Breath-by-breath gas analysis
			\\ - 12-lead ECG monitoring
			\\ - Initial 3-minute rest for baseline
			\\ - Incremental load to volitional fatigue

		\column[c]{0.4\textwidth}

			\includegraphics[height=0.4\textheight]{../Figures/cpet-zan}

	\end{columns}
	}
\end{frame}

@@

\begin{frame}
	\frametitle{Aims of Thesis}
	\begin{enumerate}
		\item To evaluate the clinical utility of preoperative CPET in predicting postoperative adverse events after pancreaticoduodenectomy.
\pause
		\item To examine the patient factors that are related to cardiopulmonary exercise physiology with particular attention to the effect of obstructive jaundice and body composition.
\pause
		\item To examine the effect of preoperative systemic inflammation and poor aerobic capacity on the magnitude of the post-operative systemic inflammatory response.
\pause
		\item To examine the value of serial daily postoperative markers of systemic inflammatory response in the prediction of post-operative complications.
	\end{enumerate}
\end{frame}

%--CHAPTER 2----CHAPTER 2----CHAPTER 2----CHAPTER 2----CHAPTER 2----CHAPTER 2----CHAPTER 2--
\section[Chapter 2]{CPET and Postoperative Outcomes}
\begin{frame}
	\frametitle{CPET and Postoperative Outcomes}
	\framesubtitle{Does CPET have a role in risk stratification of patients?}
	\begin{block}{\textbf{Complications}}
		Low $\dot{V}_{O_2}$AT associated with pancreatic fistula (16\% vs 35\%) \\
		Not associated with cardio-respiratory complications or mortality
	\end{block}
\pause	
	\begin{block}{\textbf{Length of stay}}
		Low $\dot{V}_{O_2}$AT associated with longer hospital stay (14 vs 20 days)\\
		Not associated with critical care stay or admissions
	\end{block}
\pause	
	\begin{block}{\textbf{Adjuvant Therapy}}
		Low $\dot{V}_{O_2}$AT associated with non-progression to adjuvant therapy after surgery.
	\end{block}
\end{frame}

@@

		\column{0.6\textwidth}
\pause
			{\footnotesize 	Why is $\dot{V}_{O_2}$AT so low in patients undergoing pancreatic surgery?\\
				\medskip

				What factors are associated with a low $\dot{V}_{O_2}$AT?\\
				\medskip
\pause
				Why was $\dot{V}_{O_2}$AT not associated with cardio-respiratory complications?\\
				\medskip

				Does $\dot{V}_{O_2}$AT truly reflect cardiopulmonary function in this cohort of patients?\\	
			}

@@

		\begin{figure}
			\includegraphics[width=\textwidth]{jaundice_vs_patient_factors}
		\end{figure}
		\pause		
		\column[c]{0.5\textwidth}		
		{\scriptsize			
			Severity of systemic inflammation proportionate to severity of jaundice. \\
			\medskip
			Obstructive jaundice associated with electrolyte abnormalities.\\
			\medskip
			Jaundiced patients were more likely to be anaemic, with lower haematocrit and lower mean corpuscular volume.\\
			\medskip
			No significant difference in the preoperative renal function, prothrombin time and white cell count between jaundiced and non-jaundiced patients.\\
		}
	\end{columns}
\end{frame}

\begin{frame}
	\frametitle{What preop factors were associated with low $\dot{V}_{O_2}$?}
	\framesubtitle{Multivariate binary logistic regression analysis}
	Factors associated with $\dot{V}_{O_2}$AT $<$ 10 ml/kg/min
	\begin{table}
		\begin{tabular}{l| l l l}
			& HR   & 95\% CI    & p     \\ \hline
			Female sex & 3.75 & 1.57-8.95  & 0.003 \\
			BMI$>$25   & 3.65 & 1.61-8.26  & 0.002 \\
			Cancer     & 4.02 & 1.33-12.16 & 0.014 \\
			CRP$>$10   & 2.98 & 1.29-6.86  & 0.010
		\end{tabular}
	\end{table}
	\hrule
\pause
	\medskip
	Factors associated with $\dot{V}_{O_2}$Peak $<$ 16ml/kg/min
	\begin{table}
		\begin{tabular}{l| l l l}
			& HR   & 95\% CI   & p        \\ \hline
			Female sex  & 7.57 & 3.09-18.5 & $<$0.001 \\
			BMI$>$25    & 2.57 & 1.18-5.63 & 0.018    \\
			Haemoglobin & 2.43 & 1.05-5.63 & 0.038
		\end{tabular}
	\end{table}
\end{frame}

@@

	\vfill
\pause	
	For example, if the absolute $\dot{V}_{O_2}$AT was 0.84 litres/min, the corrected $\dot{V}_{O_2}$AT (ml/kg/min) would vary with patient weight as follows:

@@

	\begin{figure}
		\includegraphics[height=0.4\textheight]{../Figures/bc_scatter_VO2_TAT}
	\end{figure}

	\begin{figure}
		\includegraphics[height=0.45\textheight]{../Figures/bc_scatter_VO2_TAT_elbm}
	\end{figure}

@@

@@

\begin{frame}
	\frametitle{Impact of perioperative systemic inflammation?}
	\framesubtitle{Inflammation before and after surgery affects short and long term outcomes}
	\begin{itemize}
		\item Perioperative systemic inflammation affects short and long-term outcomes
		\item Elevated preoperative systemic inflammation is associated with postoperative complications in colorectal surgery, after oesophagectomy, liver resections and cardiac surgery. 
		\item Early exaggerated postoperative systemic inflammatory response is also associated with increased infective complications
		\item In patients undergoing pancreatic surgery, preoperative inflammatory state is further influenced by:
		\begin{itemize}
			\item Obstructive jaundice
			\item Cholangitis
			\item ERCP and stenting
			\item Post-ERCP / EUS related complications
		\end{itemize}
		\item Obesity is associated with chronic inflammation
		\item Impact of aerobic capacity on postoperative inflammation not studied before
	\end{itemize}
\end{frame}

@@

\section{Case study}
\begin{frame}
	\frametitle{A case study}
	\framesubtitle{How does this work influence decision-making?}
	\begin{itemize}
		\item 63 yr old female with BMI of 35 presents with painless OJ
		\item PMH - HTN, T2DM, ex-smoker
		\item Bloods: Bil 180, CRP 22, Alb 30
		\item Imaging: Resectable tumour of head of pancreas
		\medskip \hrule \medskip
\pause
		\item Should this patient undergo CPET?
		\item Should this patient undergo biliary drainage?
		\item Should we wait until resolution of jaundice before CPET?
		\item How can the CPET results be used to influence her care?
		\item How can the postoperative inflammatory markers be used to manage her care?
	\end{itemize}
\end{frame}

\begin{frame}
	\frametitle{A case study}
	\begin{description}
	{\scriptsize 	
		\item [Chapter 2] CPET may allow identifying if she is at increased risk of complications, prolonged hospital stay and non-receipt of adjuvant therapy. 
					\\ CPET should not be used on its own to decline surgery. 
					\hrule
\pause
		\item [Chapter 3] CPET can be performed in the presence of jaundice. 
					\\ Jaundice does not adversely affect dynamic cardiopulmonary response to and this cannot be a justification for biliary drainage. 
					\\ Taken together with evidence from trials on preperative biliary drainage, she should proceed direct to surgery.
						\hrule
\pause
		\item [Chapter 4] Interpret low $\dot{V}_{O_2}$ in her case with caution. 
					\\ Her aerobic capacity may be better than that suggested by her results. 
					\\ Beware of \textit{spurious correlation} with total body weight.
						\hrule
\pause
		\item [Chapter 5] Jaundice and hypo-albuminemia attenuate post-op inflammatory response.
					\\ However, this does not translate into increased complications. 
					\\ Aerobic capacity does not affect post-op inflammatory response.
						\hrule
\pause
		\item [Chapter 6] If her drain amylase on D3 is normal and D3 CRP$<$178 or D4 CRP$<$125, she is likely to have an uneventful postoperative course and must continue on 'enhanced-recovery' pathway.
				\\ If she develops a POPF, the early inflammatory response does not predict outcome.
	}
	\end{description}
\end{frame}
